---
title: "Explore Airbnb listings in Paris, France, as at 04 March 2024"
author: 
  - Yimiao Yuan
thanks: "Code and data are available at: https://github.com/YimiaoYuan09/Airbnb_EDA_Paris"
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# install packages and load library
# install.packages("arrow")
# install.packages("mice")
# install.packages("naniar")

library(arrow)
library(janitor)
library(knitr)
library(lubridate)
library(mice)
library(modelsummary)
library(naniar)
library(tidyverse)
library(ggplot2)
```

# Introduction
This report analyzes the Airbnb listings in Paris, France, as at 04 March 2024 using R (@citeR). The dataset is read from the Inside Airbnb (@citeWeb) website, and cleaned and explored using `tidyverse` (@citeTidyverse), `janitor` (@citeJanitor), `knitr` (@citeKnitr), `lubridate` (@citeLubridate), `mice` (@citeMice), `modelsummary` (@citeModelsummary), `ggplot2` (@citeGgplot2) and `naniar` (@citeNaniar). This report will explore distribution and properties of different variables and analyze the relationship between them.

```{r}
#| include: false
#| warning: false
#| message: false

# read in the raw data
airbnb_raw <- read_csv(here::here("inputs", "data", "airbnb_raw_data.csv"))
```

```{r}
#| include: false
#| warning: false
#| message: false

# exploratory: create a parquet file with selected variables
airbnb_select <-
  airbnb_raw |>
  select(
    host_id,
    host_response_time,
    host_is_superhost,
    host_total_listings_count,
    neighbourhood_cleansed,
    bathrooms,
    bedrooms,
    price,
    number_of_reviews,
    review_scores_rating,
    review_scores_accuracy,
    review_scores_value
  )

rm(airbnb_raw)
```

```{r}
#| include: false
#| warning: false
#| message: false

# save the parquet file
write_parquet(
  x = airbnb_select, 
  sink = 
    "../../inputs/data/2024-03-04-paris-airbnblistings-select_variables.parquet"
  )
```

# Distribution and Properties
The raw dataset is cleaned and used to create a parquet file with selected variables for exploratory purposes. I first look at the price of the Airbnb and plot the distribution in @fig-price. There are outliers on regular scale, so I also use a log scale to plot the distribution. From the graph, we can see that most properties have have low prices, number of properties decreases as the price increases.

```{r}
#| include: false
#| warning: false
#| message: false
airbnb_select$price |>
  head()
```

```{r}
#| include: false
#| warning: false
#| message: false
# get character from price
airbnb_select$price |>
  str_split("") |>
  unlist() |>
  unique()
```

```{r}
#| include: false
#| warning: false
#| message: false
# , value
airbnb_select |>
  select(price) |>
  filter(str_detect(price, ","))
```

```{r}
#| include: false
#| warning: false
#| message: false
# remove $
airbnb_select <-
  airbnb_select |>
  mutate(
    price = str_remove_all(price, "[\\$,]"),
    price = as.integer(price)
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-price
#| fig-cap: "Distribution of prices of Paris Airbnb rentals in March 2024"
#| fig-subcap: ["Distribution of prices","Using the log scale for prices more than $1,000"]
#| layout-ncol: 2

# distribution of price
airbnb_select |>
  ggplot(aes(x = price)) +
  geom_histogram(binwidth = 10) +
  theme_classic() +
  labs(
    x = "Price per night",
    y = "Number of properties"
  )

# distribution of price on log scale
airbnb_select |>
  filter(price > 1000) |>
  ggplot(aes(x = price)) +
  geom_histogram(binwidth = 10) +
  theme_classic() +
  labs(
    x = "Price per night",
    y = "Number of properties"
  ) +
  scale_y_log10()
```
Then I turned my focus on price that less than \$1000, and the distribution contains more detailed. @fig-price-detail-1 shows that there are some bunching in the price, so I choose to take a further look at prices between \$90 and \$250. @fig-price-detail-2 also shows some bunching in the graph, which might due to the reason that prices around numbers ending in zero or nine.
```{r}
#| echo: false
#| message: false
#| eval: true
#| warning: false
#| label: fig-price-detail
#| fig-cap: "Distribution of prices of Paris Airbnb rentals in March 2024"
#| fig-subcap: ["Prices less than $1,000 suggest some bunching","Prices between $90 and $250 illustrate the bunching more clearly"]
#| layout-ncol: 2

# focus on price < 1000
airbnb_select |>
  filter(price < 1000) |>
  ggplot(aes(x = price)) +
  geom_histogram(binwidth = 10) +
  theme_classic() +
  labs(
    x = "Price per night",
    y = "Number of properties"
  )

airbnb_select |>
  filter(price > 90) |>
  filter(price < 250) |>
  ggplot(aes(x = price)) +
  geom_histogram(binwidth = 1) +
  theme_classic() +
  labs(
    x = "Price per night",
    y = "Number of properties"
  )
```

```{r}
#| include: false
#| warning: false
#| message: false

# remove price > 1000
airbnb_data_less_1000 <-
  airbnb_select |>
  filter(price < 1000)
```

```{r}
#| include: false
#| warning: false
#| message: false
# NA value
airbnb_data_less_1000 |>
  filter(is.na(host_is_superhost))
```

```{r}
#| include: false
#| warning: false
#| message: false

# remove NA value
# create binary value: 0/1
airbnb_data_no_superhost_na <-
  airbnb_data_less_1000 |>
  filter(!is.na(host_is_superhost)) |>
  mutate(
    host_is_superhost_binary =
      as.numeric(host_is_superhost)
  )
```

```{r}
#| include: false
#| warning: false
#| message: false
# NA in review scores rating
airbnb_data_no_superhost_na |>
  filter(is.na(review_scores_rating)) |>
  nrow()
```

```{r}
#| include: false
#| warning: false
#| message: false
airbnb_data_no_superhost_na |>
  filter(is.na(review_scores_rating)) |>
  select(number_of_reviews) |>
  table()
```
Next I look at the superhosts in the Airbnb. After removing the NA value in superhost, I plot the distribution of review scores for all superhosts. @fig-review-scores-rating shows that most of the superhosts' properties have 5 stars, they rarely get 1 or 2 stars.
```{r}
#| echo: false
#| message: false
#| eval: true
#| warning: false
#| label: fig-review-scores-rating
#| fig-cap: "Distribution of review scores for Paris Airbnb rentals in March 2024"

# no NA in review scores rating
airbnb_data_no_superhost_na |>
  filter(!is.na(review_scores_rating)) |>
  ggplot(aes(x = review_scores_rating)) +
  geom_histogram(binwidth = 1) +
  theme_classic() +
  labs(
    x = "Average review score",
    y = "Number of properties"
  )
```

```{r}
#| include: false
#| warning: false
#| message: false

# remove NA in review scores rating
airbnb_data_has_reviews <-
  airbnb_data_no_superhost_na |>
  filter(!is.na(review_scores_rating))
```

```{r}
#| include: false
#| warning: false
#| message: false
# host response time
airbnb_data_has_reviews |>
  count(host_response_time)
```

There are a lot of NA values in the review scores rating since they do not have enough reviews. For this report, I just remove NA values in review scores rating. 

```{r}
#| include: false
#| warning: false
#| message: false

# recode N/A to be actual NAs
# class change to factor
airbnb_data_has_reviews <-
  airbnb_data_has_reviews |>
  mutate(
    host_response_time = if_else(
      host_response_time == "N/A",
      NA_character_,
      host_response_time
    ),
    host_response_time = factor(host_response_time)
  )
```

I also interested in the host response time, which is how quickly a host responds to an inquiry. There are also a lot of NAs in this variable, so I want to see that if there is a relationship with the review score. @fig-host-response-time shows that most of properties get a rate over 4 stars. However, `ggplot2` drops a lot of missing values, so I use geom_miss_point() from `naniar` to include them in the graph (@fig-host-response-time-missing).

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-host-response-time
#| fig-cap: "Distribution of review scores for properties with NA response time, for Paris Airbnb rentals in March 2024"

# host response time NA
# relationship with review scores rating
airbnb_data_has_reviews |>
  filter(is.na(host_response_time)) |>
  ggplot(aes(x = review_scores_rating)) +
  geom_histogram(binwidth = 1) +
  theme_classic() +
  labs(
    x = "Average review score",
    y = "Number of properties"
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-host-response-time-missing
#| fig-cap: "Missing values in Paris Airbnb data, by host response time"

# include missing data
airbnb_data_has_reviews |>
  ggplot(aes(
    x = host_response_time,
    y = review_scores_accuracy
  )) +
  geom_miss_point() +
  labs(
    x = "Host response time",
    y = "Review score accuracy",
    color = "Is missing?"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
#| include: false
#| warning: false
#| message: false
# remove NA in host_response_time
# superhost, has revies, has response time
airbnb_select <-
  airbnb_data_has_reviews |>
  filter(!is.na(host_response_time))
```

I also interested in how many properties a host has on Airbnb. From @fig-host-properties, we can see that most hosts have 1 properties, some hosts have less than 10 properties. Few of them have more than 100 properties, which is a little bit strange.
```{r}
#| warning: false
#| message: false
#| eval: true
#| echo: false
#| label: fig-host-properties
#| fig-cap: "Distribution of the number of properties a host has on Airbnb, for Paris Airbnb rentals in March 2024"

# how many properties a host has on Airbnb
airbnb_select |>
  ggplot(aes(x = host_total_listings_count)) +
  geom_histogram() +
  scale_x_log10() +
  labs(
    x = "Total number of listings, by host",
    y = "Number of hosts"
  )
```

```{r}
#| include: false
#| warning: false
#| message: false
# host with number of listings > 500
airbnb_select |>
  filter(host_total_listings_count >= 500) |>
  head()
```

```{r}
#| include: false
#| warning: false
#| message: false
# focus on host with only 1 property
airbnb_select <-
  airbnb_select |>
  add_count(host_id) |>
  filter(n == 1) |>
  select(-n)
```

\newpage

# Relationship between variables
After explore distribution of individual variables, I also interested in relationship between variables. @fig-price-review shows the relationship between price and review and whether a host is a superhost, for properties with more than 1 review. From the graph, we can see that most properties have price less than /$250 and gain a 4 or 5 stars review. If the host is a superhost, the properties are more likely have higher price and have more 5 stars review. As the price increases, the average review score also increases. 
```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-price-review
#| fig-cap: "Relationship between price and review and whether a host is a superhost, for Paris Airbnb rentals in March 2024"

# more then 1 review
airbnb_select |>
  filter(number_of_reviews > 1) |>
  ggplot(aes(x = price, y = review_scores_rating, 
             color = host_is_superhost)) +
  geom_point(size = 1, alpha = 0.1) +
  theme_classic() +
  labs(
    x = "Price per night",
    y = "Average review score",
    color = "Superhost"
  ) +
  scale_color_brewer(palette = "Set1")
```

```{r}
#| include: false
#| warning: false
#| message: false
# proportion of superhost
airbnb_select |>
  count(host_is_superhost) |>
  mutate(
    proportion = n / sum(n),
    proportion = round(proportion, digits = 2)
  )
```

```{r}
#| include: false
#| warning: false
#| message: false
# host response time, by superhost
airbnb_select |>
  tabyl(host_response_time, host_is_superhost) |>
  adorn_percentages("col") |>
  adorn_pct_formatting(digits = 0) |>
  adorn_ns() |>
  adorn_title()
```

```{r}
#| include: false
#| warning: false
#| message: false
# neighbourhood
airbnb_select |>
  tabyl(neighbourhood_cleansed) |>
  adorn_pct_formatting() |>
  arrange(-n) |>
  filter(n > 100) |>
  adorn_totals("row") |>
  head()
```

\newpage

# Model
A model is run on the dataset to gain a better understanding of relationships between multiple variables. This model is going to predict whether someone is a superhost, using host response time and review scores rating. Since the output is a binary value, I use the logistic regression for this model. The result shows that each variable is positively associated with the probability of being a superhost.

```{r}
#| include: false
#| warning: false
#| message: false
# forecast whether someone is a superhost
# logistic regression
# affected by faster responses and better reviews
logistic_reg_superhost_response_review <-
  glm(
    host_is_superhost ~
      host_response_time +
      review_scores_rating,
    data = airbnb_select,
    family = binomial
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
modelsummary(logistic_reg_superhost_response_review)
```

```{r}
#| include: false
#| warning: false
#| message: false
# save analysis data
write_parquet(
  x = airbnb_select, 
  sink = "../../outputs/data/2024-03-04-paris-airbnblistings-analysis_dataset.parquet"
  )
```


\newpage


# References
